---
title: "Project3_Analysis"
author: "Hantz Angrand"
date: "October 22, 2018"
output: pdf_document
---
#Load required packages
```{r}
library(plyr)
```

```{r}
library(ggplot2)
```

```{r}
library(dplyr)
```
```{r}
library(tidyr)

```

#Step 1
Load the data
```{r}
indeed_url<-url('https://raw.githubusercontent.com/mlforsachid/Data607-Project3/master/Data/Indeed_Job_Search_Results.csv')
read_indeed_url<-read.csv(indeed_url,header = TRUE, stringsAsFactors = TRUE, sep = ',')
head(read_indeed_url)
```


#Step2 Tidying the data
```{r}
#Removing the radius column
read_indeed_url<-read_indeed_url[,-5]
head(read_indeed_url)

```

```{r}
#Rename the columns
names(read_indeed_url)<-c('Source', 'Job Title','Skills','City','url','Count')
head(read_indeed_url)
```

Removing NA value in the dataset
```{r}
read_indeed_url<-na.omit(read_indeed_url)
head(read_indeed_url)
```



#Aggregate the skills to get the frequency
```{r}
indeed_skillaggr<-aggregate(read_indeed_url$Count,by=list(Category=read_indeed_url$Skills), FUN=sum)
indeed_skillaggr
```
Jobs by skills
```{r}
skills_count<-read_indeed_url %>%
  group_by(Skills) %>%
  summarise(Total=sum(Count)) %>%
  arrange(desc(Total))

skills_count
```

Jobs opening by City
```{r}
skills_city<-read_indeed_url %>%
  group_by(Skills,City) %>%
  summarise(Total=sum(Count)) %>%
  arrange(desc(Total))

skills_city
```



#Step3 Data Visualization
```{r}
plots_top<-tail(skills_count,20)

ggplot(plots_top, aes(plots_top$Skills, plots_top$Total)) + geom_bar(stat="identity")
```



```{r}
top10_skills<-skills_city[1:10,]
ggplot(top10_skills, aes(x=Skills, y=Total, colour= City, size = Total)) + geom_point()
```

```{r}
library(wordcloud)

wordcloud(skills_count$Skills,skills_count$Total, random.order=FALSE, colors=brewer.pal(8,"Dark2"))
```


#Step4 Conclusion
The top 5 skills are oriented towards Big Data, Python, R are the word most used in the dataset.  Our finding show a few skills underlie the field of data science.Certainly mathematics and Statistics should play a role in our study but they are not part of our dataset.  We cannot conclude definitely about on demand skills since some other skills are missing in the dataset.

